---
title: "vivid_Range_of_Responses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vivid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 7
)
```

This vignette is intended to be used as a supplementary tools to the introduction to the `vivid` package vignette. The goal of this vignette is to show how `vivid` handles a range of different response variables and to show a brief guide on how to create visualisations from the package to display varible importance and variable interactions (VIVI). To begin we show a standard response from a random forest regression model. 

First, we load the required packages. `vivid` to create the visualizations and `mlr3` and `mlr3learners` will be used to create our models. 

```{r, message=FALSE, warnings = F}
library(vivid)
library(mlr3)  
library(mlr3learners)
```


## Regression with a continuous response

For this we will simulate some data from the Friedman benchmark problem 1 using the `genFriedman()` function from `vivid`, with 9 features and 300 observations. The output is created according to:

<center>
$$y = 10 sin(π x1 x2) + 20 (x3 - 0.5)^2 + 10 x4 + 5 x5 + e$$
</center>


Create  the data:
```{r, messages = FALSE}
set.seed(1701)
myData <- genFriedman(noFeatures = 9, noSamples = 300, sigma = 1, bins = NULL, seed = NULL)
```

Create an `mlr3` `ranger` random forest model:

```{r}
set.seed(1701)
taskCont  <- TaskRegr$new(id = "Friedman", backend = myData, target = "y")
set.seed(1701)
lrnCont <- lrn("regr.ranger", importance = "permutation")
set.seed(1701)
modCont <- lrnCont$train(taskCont)
```

Create a matrix to be supplied to the plotting functions.

```{r, message = F, warning = F}
set.seed(1701)
myMatrix <- vividMatrix(task = taskCont, model = modCont, gridSize = 100)

```

Create a heatmap of the results

```{r, fig.width = 7, fig.height = 7}
plot(myMatrix, type = "heatMap")
```
<center>
Fig 1.0: *Heat-map plot displaying 2-way interaction strength in blue and individual variable importance on the diagonal in red.*
</center>

Figure 1.0 shows a heatmap of VIVI values on the Friedman data. We can clearly see the variables $x_1$ to $x_5$ are the only important variables for predeicting $y$, with $x_4$ being the most important. We can also seee a strong interaction between $x_1$ and $x_2$ 
Create a partial dependence pairs plot (GPDP)

```{r}
ggpdpPairs(taskCont, modCont, gridsize = 20)
```

Create a zenplot partial dependence plot (ZPDP)

```{r}
pdpZenplot(taskCont, modCont, gridsize = 20)
```


## Numerical Binary Response 

Here we can use `genFriedman` again to create data with a binary response. To do this we set the number of bins to be equal to 2. This will create (roughly) equal sized bins to split the response into. Setting a value greater than 1 essentially turns this into a classification problem where bins determines the number of classes.


Create  the data:
```{r, messages = FALSE}
set.seed(1701)
myDataBin <- genFriedman(noFeatures = 9, noSamples = 300, sigma = 1, bins = 2, seed = NULL)
myDataBin$y <- as.numeric(myDataBin$y)
```

Take a quick look at our data to see that the response has been split into 2 separate classes.

```{r}
head(myDataBin)
```

Create an `mlr3` `ranger` random forest model:

```{r}
set.seed(1701)
taskBin  <- TaskRegr$new(id = "Friedman", backend = myDataBin, target = "y")
set.seed(1701)
lrnBin <- lrn("regr.ranger", importance = "impurity")
set.seed(1701)
modBin <- lrnCont$train(taskCont)
```

Create a matrix to be supplied to the plotting functions.

```{r, message = F, warning = F}
set.seed(1701)
myMatrixBin <- vividMatrix(task = taskBin, model = modBin, gridSize = 100)

```

Create a heatmap of the results

```{r, fig.width = 7, fig.height = 7}
plot(myMatrixBin, type = "heatMap")
```

Create a partial dependence pairs plot (PDPP)

```{r}
ggpdpPairs(taskBin, modBin, gridsize = 20)
```

Create a zenplot partial dependence plot (ZPDP)

```{r}
pdpZenplot(taskBin, modBin, gridsize = 20)
```


## Multiclass Categorical Response

For this we will use the iris dataset to show how to use `vivd` when there is a multiclass categorical response.

To begin, we create our `mlr3` model.

```{r}
# mlr3 TASK
set.seed(1701)
ir_T <- TaskClassif$new(id = "iris", backend = iris, target = "Species")
# learner
set.seed(1701)
lrn <- lrn("classif.ranger", importance = 'impurity')
# model
set.seed(1701)
ir_M <- lrn$train(ir_T)
```

To create our vivid matrix of values, we have to specify the main variable of interest by use of the `main` argument. In the example below, we choose to look at the species "virginica".

```{r}
myMatrixMCR <- vividMatrix(ir_T, ir_M, gridsize = 100,  main = "virginica")
```

Plot the results:
```{r, fig.width = 7, fig.height = 5}
plot(myMatrixMCR, type = 'heatMap', title = "virginica")
```

Create a partial dependence pairs plot (PDPP)

```{r}
ggpdpPairs(ir_T, ir_M, gridsize = 20)
```

Create a zenplot partial dependence plot (ZPDP)

```{r}
pdpZenplot(ir_T, ir_M, gridsize = 20)
```
