---
title: "Variable Importance and Variable Interaction Displays (vivid)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vivid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 7
)
```

```{r setup, message=FALSE, echo = FALSE}
library(vivid)
```

## Introduction

This package was designed to help a user to easily distinguish which
variables in a model are important and which variables interact with
each other. It does this by giving the user different plotting options.
These include a heatmap style plot that displays 2-way interactions and
individual variable importance. Also, a network style plot where the
size of a node represents variable importance (the bigger the node, the
more important the variable) and the edge weight represents the 2-way
interaction strength. Also included is an option to display partial
dependence plots that utilize an Eularian tour to help the user to
easily identify which variables interact the most. The interaction is
calculated using *Friedman’s H-Statistic*\[1\].

## Data used in this vignette:

The data used in the following examples is the *air quality* data which
is the daily air quality measurements in New York, May to September
1973\[2\]. The data contains 6 variables which are described as follows:

  - Ozone: Mean ozone in parts per billion from 1300 to 1500 hours at
    Roosevelt Island.

  - Solar.R: Solar radiation in Langleys in the frequency band 4000–7700
    Angstroms from 0800 to 1200 hours at Central Park.

  - Wind: Average wind speed in miles per hour at 0700 and 1000 hours at
    LaGuardia Airport.

  - Temp: Maximum daily temperature in degrees Fahrenheit at La Guardia
    Airport.

  - Month: The month the data was recorded (takes a value between 1-12).

  - Day: The day the the data was recorded (takes a value between 1-7).

In all the following examples, *Ozone* is used as the response variable.

## Commands:

# intHeatmap()

-----

*Create a Heatmap style plot displaying Variable Importance and Variable
Interaction*

-----

**Description**

Plots a heatmap of the interaction strength with variable importance on
the off-diagonal

**Usage**

intHeatmap(task, model, method, interact)

**Arguments**

  - task - Task created from the `mlr` package (either regression or
    classification).

  - model - Any machine learning model.

  - method - A list of variable importance methods to be set by the
    user. These can include any of the importance methods contained
    within the `mlr` package. The default is method = randomForest.

  - interact - If interact = TRUE, an interactive plotly object is
    displayed where the user can hover the mouse icon over any part of
    the plot and a text box will appear displaying information related
    to the selected variables.

  - intLow & intHigh - The interaction strength values are displayed
    using a colour gradient. `intLow` is a colour, declared by the user,
    which is assigned to low interaction strength values. `intHigh` is
    the colour assigned to high interaction strength values.

  - impLow & impHigh - Similarly, `impLow` and `impHigh` represent the
    low to high values of the colour gradient used to display the
    individual variable
importance.

**Examples**

Load in the data:
```{r, messages = FALSE}
aq <- data.frame(airquality)
aq <- na.omit(aq)
```

Run an `mlr` random forest smodel:
```{r}
library(mlr)
library(randomForest)
aqRgrTask  <- makeRegrTask(data = aq, target = "Ozone")
aqRegrLrn <- makeLearner("regr.randomForest")
aqMod <- train(aqRegrLrn, aqRgrTask)
```

To call the plot we use the `intHeatmap()` command as follows:

```{r}
intHeatmap(task = aqRgrTask, model = aqMod, method = "randomForest_importance", interact = F)
```
